---
title: "Feb 21 Exercises"
author: "Matthew D. CIaramitaro"
date: "February 21, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Exercises 10.5
<h1>10.5.1</h1>
We can tell if a dataframe is a tibble by printing it and seeing if it follows the print behavior of a tibble. If it prints more than 10 rows it is not a tibble. However, this method doesn't work for all cases.The function `is.tibble` below would work for all cases.
```{r}
df <- mtcars
tb <- as.tibble(mtcars)
print(is.tibble(df))
print(is.tibble(tb))
```
<h1>10.5.2</h1>
```{r}
#first with the data.frame
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]
#now with the tibble
df <- as.tibble(data.frame(abc = 1, xyz = "a"))
df$x
df[, "xyz"]
df[, c("abc", "xyz")]
```
The tibble doesn't have access to the same indexing keys as the data frame. We can no longer use the old methods to obtain the data by row/column name.
Dataframes can be frustrating because unlike tibbles, they do not inform you that you accessed a column that doesn't exist and do partial matches like `df$x` yielding `df$xyz` in the above.

<h1>10.5.3</h1>
We can use piping or normal indexing to extract 
```{r}
var <- "mpg"
#let's use tb from before
tb[[var]]
tb %>% .[[var]]
```
<h1>10.5.4</h1>
Using the following dataframe we will answer some questions
```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)
```
<h2>10.5.4.1</h2>
Extract the variable 1
```{r}
annoying %>% .[["1"]]
```
<h2>10.5.4.2</h2>
```{r}
plot(x = annoying %>% .[["1"]], y = annoying %>% .[["2"]], main="1 vs 2 scatterplot", type="p", xlab = "1", ylab = "2")
```
<h2>10.5.4.3</h2>
We use the add column function
```{r}
annoying <- add_column(annoying, "3"= annoying[["2"]]/annoying[["1"]])
annoying
```
<h2>10.5.4.4</h2>
```{r}
annoying <- rename(annoying, "one" = "1", "two" = "2", "three" = "3")
annoying
```
<h1>10.5.5</h1>
This tidies the data, giving it two columns, one for variable and one for data.

<h1>10.5.6</h1>
`options(tibble.width = n)` will make the number of columns printed n. If n is `Inf` all columns will be printed.

##Exercises 12.6.1


